### 2014-01-26 Papers_SVN/papers/unpublished/2010-energy_accounting_extended-arliones/files/1_intro.tex
Versionando o meu mestrado.

\fig{sampled_real_discharge}{Sampled discharge plotted against real discharge
for a CR2 series Panasonic lithium
battery~\cite{Panasonic:Lithium:2006}.}{scale=.333}

To cope with this requirement mobile systems provide means to measure battery
voltage by which it is possible to infer battery charge through an approximate
discharge model for a given battery. This approach, however, brings limitations
to the task of estimating battery charge.
%which adds to the \emph{good} voltage stability of Lithium-based batteries.
Among the reported problems~\cite{Mundra:2008,Penella:2010} there is one of
special interest for the task of precisely estimating battery charge on
embedded systems: the low accuracy and long response time of voltage-based
battery state-of-charge models. This problem is related to the diminished
precision of such voltage measurements implied by low resolution
analog-to-digital converters and to the oscillation on the battery voltage-levels
due to load variations. It can be easily illustrated by
Fig.~\ref{fig:sampled_real_discharge}. The figure shows that during most of
system lifetime the response time in terms of battery voltage varies greatly.
For instance, if an energy-related decision lowers system utilization on day 50,
this decision may last until the next expected voltage drop, around day 100,
making an unique decision stay in effect for $14\%$ of system lifetime. We
can also see that the scheduler misses the opportunity of raising system utilization
because its monitor is not able to rapidly inform that the system is not using
the expected amount of energy. Instead of gradually raising system utilization
and keeping it higher during a longer period of time, it only acts by the end of
a given period, causing bursts in system utilization for short periods of time.

In this paper we propose a software-based energy consumption accounting scheme.
We rely on the fact that energy consumption is a cross-cutting concern,
orthogonal to all system components~\cite{Lohmann:2005}, to enable accurate
monitoring of energy consumption in a component-based operating system for
embedded systems. This scheme monitors the execution of operating system
functionalities. In order to make this approach feasible, we derived three
different profiles for energy accounting which can be applied to devices with
different operational behavior. We evaluate the effect of our approach on system
performance. This paper focus exclusively in the proposed model. An
implementation of the system was also performed for the
\epos~platform~\cite{Project:EPOS:2010} by extending \epos' power management
scheme~\cite{Froehlich:2011}, altought the implementation details where kept out
of this paper due to length limitations.

The remaining of this paper is organized as follows.
Section~\ref{sec:related} presents an overview of related work.
Section~\ref{sec:account} describes the software-based energy accounting
approach proposed in this paper, analyzing it through a didactic case-study.
% Section~\ref{sec:impl_epos} describes the implementation of this approach in the
% \epos~Project.
Section~\ref{sec:case} presents a case-study of the proposed approach on a
wireless sensor network system.
Section~\ref{sec:concl} summarizes this paper and gives some insights on future
work.

>>>>

\fig{sampled_real_discharge}{Sampled discharge plotted against real discharge
for a CR2 series Panasonic lithium
battery~\cite{Panasonic:Lithium:2006}.}{scale=.37}

To cope with this requirement mobile systems provide means to measure battery
voltage by which it is possible to infer battery charge through an approximate
discharge model for a given battery. This approach, however, brings limitations
to the task of estimating battery charge.
%which adds to the \emph{good} voltage stability of Lithium-based batteries.
Among the reported problems~\cite{Mundra:2008,Penella:2010} there is one of
special interest for the task of precisely estimating battery charge on
embedded systems: the low accuracy and long response time of voltage-based
battery state-of-charge models. This problem is related to the diminished
precision of such voltage measurements implied by low resolution
analog-to-digital converters and to the oscillation on the battery voltage-levels
due to variations on the applied charge. It can be easily illustrated by
Figure~\ref{fig:sampled_real_discharge}. The figure shows that during most of
system lifetime the battery-voltage response time varies greatly. For instance,
if an energy-related decision lowers system utilization on day 50, this decision
may last until the next expected voltage drop, around day 100, making an unique
decision stay in effect for $14\%$ of system lifetime. We can also see that the
scheduler misses the opportunity of raising system utilization because its
monitor is not able to rapidly inform that the system is not using the expected
amount of energy. Instead of gradually raising system utilization and keeping it
higher during a longer period of time, it only acts by the end of a given
period, causing bursts in system utilization for short periods of times.

In this paper we propose a software-based energy consumption accounting scheme.
We rely on the fact that energy consumption is a cross-cutting concern,
orthogonal to all system components~\cite{Lohmann:2005}, to enable accurate
monitoring of energy consumption in a component-based operating system for
embedded systems. This scheme monitors the execution of operating system
functionalities. In order to make this approach feasible, we derived two
different profiles for energy accounting which can be applied to devices with
different operational behavior. We evaluate the effect of out approach on system
performance.

The remaining of this paper is organized as follows.
Section~\ref{sec:related} presents an overview of related work.
Section~\ref{sec:account} describes the software-based energy accounting
approach proposed in this paper.
Section~\ref{sec:impl_epos} describes the implementation of this approach in the
\epos~Project.
Section~\ref{sec:case} presents a case-study of the proposed approach on a
wireless sensor network system.
Section~\ref{sec:concl} summarizes this paper and gives some insights on future
work.
